---
title: "Penguins Size Comparisons"
author: "Alex Gaynor"
date: "11/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(effsize)
library(broom)
library(ggbeeswarm)
library(car)

```

#Part 1. T-tests

Mass comparison between male and female penguins.

```{r}
chinstrap <- penguins %>% 
  filter(species == "Chinstrap")

# histograms - we don't give a y variable. The histogram counts the things in the bins to populate the y axis.
ggplot(data = chinstrap, aes(x = body_mass_g)) +
  geom_histogram(bins = 10) +
  facet_wrap(~sex)

# qq plot - the closer to linear, the closer the sample is to being drwn from a normal distribution
ggplot(data = chinstrap, aes(sample = body_mass_g)) +
  geom_qq() +
  facet_wrap(~sex)

# Get group descriptive statistics
chinstrap %>%
  group_by(sex) %>% 
  summarize(
    mean_mass = mean(body_mass_g),
    median_mass = median(body_mass_g),
    sd_mass = sd(body_mass_g),
    sample_size = n()
  )
```

These qq distribution are about as close to linear as real world data gets.


### Get vectors f body_mass_g for femal eand male chinstrap penguins

```{r}
chinstraps_f <- chinstrap %>% 
  filter(sex == "female") %>% 
  pull(body_mass_g)

chinstraps_m <- chinstrap %>% 
  filter(sex == "male") %>% 
  pull(body_mass_g)
```

### T-tests, some different ways

1. is there a significant difference in body mass between male and femal echinstrap penguins?

```{r}
chinstrap_ttest <- t.test(chinstraps_f, chinstraps_m)
chinstrap_ttest

cohen.d(chinstraps_f, chinstraps_m)
```

Are male chinstrap pengiuns significantly larger in mass than female chinstrap penguins. This is a one-sided t-test. We only want to know one tail of the probability.

```{r}
t.test(chinstraps_m, chinstraps_f, alternative = "greater")
```

## One-way ANOVA

### 1. table of counts by species and sex

```{r}
penguin_counts <- penguins %>% 
  count(species, sex)
penguin_counts
```


create a subset of female penguins

```{r}
penguin_f <- penguins %>% 
  filter(sex == "female")

ggplot(data = penguin_f, aes(x = flipper_length_mm)) +
  geom_histogram(aes(fill = species)) +
  facet_wrap(~species, scales = "free")

ggplot(data = penguin_f, aes(sample = flipper_length_mm)) +
  geom_qq(aes(color = species)) +
  facet_wrap(~species, scales = "free")
  
```


### Summary Statistics

Starting from penguin_f, create a summary table that groups by species, and groups for the following:

- mean of slipper length
- sd
- sample size
- variance `var`
- standard error = sd/sqrt(n)

```{r}
penguin_f_summary <- penguin_f %>% 
  group_by(species) %>% 
  summarize(
    mean_flipper_length = mean(flipper_length_mm),
    sd_flipper_length = sd(flipper_length_mm),
    sample_size = n(),
    var_flipper_length = var(flipper_length_mm),
    se_flipper_length = sd(flipper_length_mm)/sqrt(n())
  )

leveneTest(flipper_length_mm ~ species, data = penguin_f)
```

### Beeswarm plot with +- 1 sd

```{r}
ggplot() +
  geom_beeswarm(data = penguin_f,
                aes(x = species, y = flipper_length_mm),
                color = "gray70",
                size = 2,
                alpha = 0.5) +
  geom_point(data = penguin_f_summary, aes(x = species, y = mean_flipper_length)) +
  geom_errorbar(data = penguin_f_summary,
                aes(x = species,
                    ymax = mean_flipper_length + sd_flipper_length, 
                    ymin= mean_flipper_length - sd_flipper_length),
                    width = 0.1)
```

### One-way ANOVA

```{r}
penguin_aov <- aov(flipper_length_mm ~ species, data = penguin_f)

summary(penguin_aov)

TukeyHSD(penguin_aov)
```

Type III Anova with the car package:
```{r}
penguin_lm <- lm(flipper_length_mm ~ species, data = penguin_f)
penguin_aov_t3 <- Anova(penguin_lm, type = "III")
penguin_aov_t3
```



